Test

### 2023-12-19 template/paper.qmd
Wrote literature review and additional results

---
title: "Test"
author: "Ian Berlin"
format:
  cmc-article-pdf:
    fontsize: 12pt
    margin-left: 1.5in
    margin-right: 1.5in
    margin-bottom: 1.2in
    margin-top: 1.2in
    keep-tex: true
    fig-pos: 'tbp'
    tbl-pos: 'tbp'
    include-in-header: header.tex
bibliography: aqrd-bibliography.bib
abstract: | 
  Recent studies have shown that wealth among the 400 richest Americans — the _Forbes_ 400 — has grown increasingly concentrated among individuals who inherited their wealth.  At the same time, these individuals tend to be more stable within the Forbes 400 over time.  One potential explination for this divergence could be that individuals who belong to dynastic fortunes are better able to collude to avoid state level estate taxes. Building off of @moretti_taxing_2023, I replicate their findings that billionaires are generally sensitive to estate taxes and move between states to avoid taxation, but find that there is no difference in the rate of movement depending on whether and invidiaul belongs to a dynasty or not.  
---

@@

This paper attempts to pull together two recently emerging strands within the literature on extreme wealth.  In @korom_enduring_2017, the authors

@@

Next, they link the *Forbes* list to state economic data, including estate tax rates, income tax rates, state GDP information, and population.  They also construct their own variable of dynastic fortunes, largely based off of the reported source by *Forbes*, combined with additional data from @kaplan_its_2013 that tracks whether an individual is the first generation, second generation *etc*. to appear on the *Forbes* list. Meanwhile, @fernholz_rising_2023 tracks every individual longitudinally on the *Forbes* list, grouping them together into dynasties. One shortcoming of this measure, however, is that they treat individuals who appear in the first year of the *Forbes* list all as seeding their own dynasty, even if they are clearly related to each other (for example, some members of the DuPont or Rockefeller families are listed as belonging to their own distinct line).  While @moretti_taxing_2023 collapse the generational information from @kaplan_its_2013 into a binary variable, it would be an fruitful area of future study to interrogate whether the results are robust to more finely grained measures of dynasty.  However, the Kaplan data only covers four years, limiting the availability of such a measure for the entire panel without building up a conceptually similar measure from scratch.  
```{=tex}
\begin{table}
\caption{\textbf{Summary Statistics (Individual-Year Data)}}
\label{tbl-one}
\input{tables/summarystats.tex}
\end{table}
```

@@

@@

![Summary View of Estate Tax Treatment by State, 1982-2017](figures/panelview.png){#fig-panel}

@@

Following @moretti_taxing_2023, I "use the elimination of the federal credit for state estate taxes in 2001 for assessing the locational sensitivity of the ultra-rich to estate taxes."  In particular, while billionaires generally paid the same in taxes regardless of location before 2001, their tax burdens varied substantially by state post 2001.  @fig-5 shows the share of billionaires living in a state that had an estate tax in 2001 each year.  The horizontal line shows the passage of the EGTRRA with the shaded region showing the phase in period.  The horizontal lines correspond to the pre-2001 and post-2001 mean, showing a drop from about 21 percent on average to roughly 17 percent. @fig-5 replicates Figure 5 from @moretti_taxing_2023.

![Share of Forbes 400 Living in a 2001 Estate Tax State, 1982-2017](figures/figure5.png){#fig-5}
To test this relationship more rigorously, I replicate Moretti and Wilson's [-@moretti_taxing_2023] difference-in-difference design that compares the change in the number of billionaires (or their wealth) prior to the 2001 credit elimination and after the elimination in estate tax states and non–estate tax states.  Note that while the estimates match the original paper precisely, Moretti and Wilson use Driscoll-Kraay (with ten-year bandwidth) standard errors, whereas I use simple clustered standard errors, thus explaining the difference.

|                         |
|:------------------------|
| ![](tables/reg1.png) |

\[**SK: The quarto code below puts in a figure into a markdown table and calls it a table. Inserting tables are more complicated than inserting figures, so a hack is to include the image of the table as I have done here. Alternatively, you can also ask R to generate tables in a .tex format and ask quarto read it in as a tex file. See Appendix.**\]

\[**SK: Do not delete the two lines of quarto code that starts with `::: {#refs}`. This produces a list of references.**\]

>>>>

---
title: "Test"
author: "Ian Berlin"
format:
  cmc-article-pdf:
    fontsize: 12pt
    margin-left: 1.5in
    margin-right: 1.5in
    margin-bottom: 1.2in
    margin-top: 1.2in
    keep-tex: true
    fig-pos: 'tbp'
    tbl-pos: 'tbp'
    include-in-header: header.tex
bibliography: aqrd-bibliography.bib
abstract: | 
  Recent studies have shown that wealth among the 400 richest Americans — the _Forbes_ 400 — has grown increasingly concentrated among individuals who inherited their wealth.  At the same time, these individuals tend to be more stable within the Forbes 400 over time.  One potential explination for this divergence could be that individuals who belong to dynastic fortunes are better able to collude to avoid state level estate taxes. Building off of @moretti_taxing_2023, I replicate their findings that billionaires are generally sensitive to estate taxes and move between states to avoid taxation, but find that there is no difference in the rate of movement depending on whether and individual belongs to a dynasty or not.  
---

@@

This paper attempts to pull together two recently emerging strands within the literature on extreme wealth.  In @korom_enduring_2017, the authors track self-made vs inherited fortunes, finding that although self-made fortunes have grown increasingly prominent within the *Forbes* 400 over time, inherited fortunes are more stable over time.  In other words, possessing an inherited fortune decreases the probability that an individual falls off the *Forbes* list in any given year, all else being equal.  At the same time, @fernholz_rising_2023 study the *Forbes* panel from 1982-2020, linking members by family dynasty.  They document a rise in wealth concentration and a decline in mobility for the wealthiest families over time.  Clearly, there is some benefit to having an inherited or dynastic fortune when trying to maintain or grow extreme wealth.  Yet the literature has done little probe potential mechanisms for this divergence.

@korom_enduring_2017 advance three potential explications: "preservation of capital through professional management of assets, tax avoidance, and the continuing control over family companies."  I focus here specifically on tax avoidance.  Although disaggregated information on tax avoidance is hard to estimate, making comparisons between family fortunes and self-made individuals difficult, we can test an extreme form of tax-avoidance: interstate migration to avoid the estate tax.  @moretti_taxing_2023 provides a valuable basis for considering estate tax avoidance among the extremely wealthy.

They first provide a valuable contribution to the literature by validating the residence locations that *Forbes* provides.  They do this in two ways: 1) by showing that the residence listed for billionaires who died during the observation period matching the location listed on their obituary 90 percent of time and 2) by showing that estate and inheritance tax revenues spike in the "correct" state following the death of a *Forbes* billionaire by an average 8.25 percent of that billionaires wealth, which is roughly half of the typical rate in ET states.  This mirrors findings from @raub_comparison_2010, which finds that billionaires report a value of wealth for tax purposes that is roughly half of the wealth Forbes reports.

@moretti_taxing_2023 then exploit changes in 2001 to the federal tax code which phased out federal credits for state estate tax liabilities, thereby introducing considerable interstate variation in total tax burden depending on location for billionaires. Using a difference-in-diffence design, they show that after 2001, billionaires on net move from ET states to non-ET states.  Furthermore, they use a triple-difference approach to consider the impact of age, showing that the effect is particularly strong for billionaires over 65.  Building off of that design, I test whether there is a higher sensitivity for dynasties over self-made individuals, thereby testing an extreme case of the tax-avoidance theory underlying the observations from @korom_enduring_2017 and @fernholz_rising_2023.

@@

```{=tex}
\begin{table}
\caption{\textbf{Summary Statistics (Individual-Year Data)}}
\label{#tbl-one}
\input{tables/summarystats.tex}
\end{table}
```
Next, they link the *Forbes* list to state economic data, including estate tax rates, income tax rates, state GDP information, and population.  They also construct their own variable of dynastic fortunes, largely based off of the reported source by *Forbes*, combined with additional data from @kaplan_its_2013 that tracks whether an individual is the first generation, second generation *etc*. to appear on the *Forbes* list. Meanwhile, @fernholz_rising_2023 tracks every individual longitudinally on the *Forbes* list, grouping them together into dynasties. One shortcoming of this measure, however, is that they treat individuals who appear in the first year of the *Forbes* list all as seeding their own dynasty, even if they are clearly related to each other (for example, some members of the DuPont or Rockefeller families are listed as belonging to their own distinct line).  While @moretti_taxing_2023 collapse the generational information from @kaplan_its_2013 into a binary variable, it would be an fruitful area of future study to interrogate whether the results are robust to more finely grained measures of dynasty.  However, the Kaplan data only covers four years, limiting the availability of such a measure for the entire panel without building up a conceptually similar measure from scratch.

@@

![Summary View of Estate Tax Treatment by State, 1982-2017](figures/panelview.png){#fig-panel width="87%"}

@@

@@

Following @moretti_taxing_2023, I "use the elimination of the federal credit for state estate taxes in 2001 for assessing the locational sensitivity of the ultra-rich to estate taxes."  In particular, while billionaires generally paid the same in taxes regardless of location before 2001, their tax burdens varied substantially by state post 2001.  @fig-five shows the share of billionaires living in a state that had an estate tax in 2001 each year.  The horizontal line shows the passage of the EGTRRA with the shaded region showing the phase in period.  The horizontal lines correspond to the pre-2001 and post-2001 mean, showing a drop from about 21 percent on average to roughly 17 percent. @fig-five replicates Figure 5 from @moretti_taxing_2023.

![Share of Forbes 400 Living in a 2001 Estate Tax State, 1982-2017](figures/figure5.png){#fig-five}
To test this relationship more rigorously, I replicate Moretti and Wilson's [-@moretti_taxing_2023] difference-in-difference design that compares the change in the number of billionaires  prior to the 2001 credit elimination and after the elimination in estate tax states and non–estate tax states.  Note that while the estimates match the original paper precisely, Moretti and Wilson use Driscoll-Kraay (with ten-year bandwidth) standard errors, whereas I use simple clustered standard errors, thus explaining the difference.  Column numbers correspond to the original column in the paper.

|                         |
|:------------------------|
| ![](tables/reg1.png) | 
: Partial replication of @moretti_taxing_2023 Table 2: Difference-in-Difference Dependent Variable: Population of Forbes 400 {#tbl-2}

Across all columns, the interaction term is negative and significant, showing that "the average estate tax state saw a drop of billionaires after 2001 compared with the average non–estate tax state" [@moretti_taxing_2023].  In column 1, which presents the simple model, the coeffcient is -2.454, which corresponds to a loss of 2.45 billionaires in estate tax states after 2001 compared to non-estate tax states.  However, states often change numerous tax laws at the same time, so migration in response to these other changes may appear to be migration in response to the estate tax.  Column 2 controls for the top marginal personal income tax (PIT) rate by state and year, and the findings are robust to this addition.  Controlling for PIT, states with estate taxes lose 2.68 billionaires after 2001 compared to non-estate tax states.

Another potential concern is that "the amenity value of non-ET states relative to ET states has increased over time for the very wealthy, leading to an increase over time in the share of the *Forbes* 400 living in non-ET states for reasons unrelated to the estate tax."  Therefore, column 3 controls for the share of high earners — individuals between the 90th and 97th percentile — in a state who would similarly benefit from amenities but do not pay the estate tax.  The negative coefficient on the interaction term (-2.32) is largely unchanged, showing that the decrease in population in ET states is not driven by improved amenities in non-ET states.  Column 5 uses per capita *Forbes* billionaires per state as the dependent variable, finding a coefficient of -6.21.  Column 8 shows that the results are robust to dropping 2002-2004 from the dataset, which is the period during which federal tax credits were being phased out.  @Moretti_taxing_2023 also show that their estimates are robust to replacing their estate tax dummy with specific tax rates, as well as subsetting by only the top 100, 200, and 300 wealthiest individuals in order to remove individuals who are more likely to drop off the *Forbes* list.

I then move to the triple-difference design.  @tbl-trip presents the results. None of the coefficients on the three-way interaction term are significant, suggesting that members of dynastic fortunes do not move to avoid the estate tax significantly more than non-dynastic individuals.  Columns 1-4 match the same first four columns of @tbl-2, controlling for PIT rate (and its interactions), high earner share, and number of billionaires per capita.  The next column uses the alternative measure of dynasty from @moretti_taxing_2023, while the following two columns show the results from combinations of the two, first coding an individual as dynsatic if either measure codes them as such and second coding and individual as dynastic if both measures code them as such.  The three-way interaction is non-significant for all three.  Finally, the last column shows that dropping the phase in period of 2002-2004 does not alter the results.

|                         |
|:------------------------|
| ![](tables/reg2.png){#fig-reg width="90%"}| 
: Triple-Difference Dependent Variable: Population of Forbes 400 {#tbl-trip}

# Conclusion
